### 2016-11-28 MSDS_Project_Markdown.Rmd
Initialize

---
title: "MSDS_Project_Jeng"
author: "Darren Jeng"
date: "November 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(stringr)
```

## Video Game Sales vs Reviews

```{r copies}
# When extracting the list, there are over 205 elements; only two of the elements contain the full list needed, the rest were just titles of the games. 
url <- "http://www.vgchartz.com/yearly/2016/USA/"
USA_YEARLY_GET <- url %>% read_html() %>% html_nodes("table") %>% html_table(fill = TRUE)
USA_Copies_Yearly_2016 <- USA_YEARLY_GET[[2]]
names(USA_Copies_Yearly_2016) <- make.names(names(USA_Copies_Yearly_2016))
names(USA_Copies_Yearly_2016)[1] <- "Position"
names(USA_Copies_Yearly_2016)[2] <- "Remove"
names(USA_Copies_Yearly_2016)[4] <- "Remove"
names(USA_Copies_Yearly_2016)[5] <- "Remove"
names(USA_Copies_Yearly_2016)[6] <- "Remove"
names(USA_Copies_Yearly_2016)[7] <- "Weeks"
names(USA_Copies_Yearly_2016)[8] <- "Yearly"
names(USA_Copies_Yearly_2016)[9] <- "Total"

USA_Copies_Yearly_2016_Cleaned <- USA_Copies_Yearly_2016[,c("Position", "Game", "Weeks", "Yearly", "Total")]

```